---
title: "Take Home Exercise 2"
description: "Mini-Challenge 3"
author: "Tai Qiu Yan"
date: "May 22, 2025"
date-modified: "last-modified"
format: html
editor: visual
execute:
  eval: true
  echo: true
  warning: false
---

# 1 Overview

# 2 Getting Started

## 2.1 Installing and loading the required libraries

Five R packages will be used. They are tidyverse, jsonlite, tidygraph, ggraph and SmartEDA.

In the code chunk below, `p_load()` of **pacman** package is used to load the R packages into R environemnt.

```{r}
pacman::p_load(tidyverse, jsonlite,
               tidygraph, ggraph,
               SmartEDA)
```

## 2.2 Importing Knowledge Graph Data

In the code chunk below, `fromJSON()` of **jsonlite** package is used to import *mc3.json* file into R and save the output object.

```{r}
MC3 <- fromJSON("data/MC3_graph.json")
MC3_schema <- fromJSON("data/MC3_schema.json")
```

### Inspecting knowledge graph structure

In the code chunk below glimpse() is used to reveal the structure of mc3 knowledge graph.

```{r}
glimpse(MC3)
```

## 3 Extracting the edges and nodes tables

`as_tibble()` of **tibble** package package is used to extract the nodes and links tibble data frames from *mc3* tibble dataframe into two separate tibble dataframes called *mc3_nodes* and *mc3_edges* respectively.

```{r}
mc3_nodes <- as_tibble(MC3$nodes)
mc3_edges <- as_tibble(MC3$edges)
```

### Initial EDA

In the code chunk below, `ExpCatViz()` of SmartEDA package is used to reveal the frequency distribution of all categorical fields in *mc3_nodes* tibble dataframe.

```{r}
ExpCatViz(data=mc3_nodes,
          col="lightblue")
```

The code chunk below uses ExpCATViz() of SmartEDA package to reveal the frequency distribution of all categorical fields in mc3_edges tibble dataframe.

```{r}
ExpCatViz(data=mc3_edges,
          col="lightblue")
```

## 4 Data Cleaning and Wrangling

### 4.1 Cleaning and wrangling nodes

Code chunk below performs the following data cleaning tasks:

-   convert values in id field into character data type,

-   exclude records with `id` value are na,

-   exclude records with similar id values,

-   exclude `thing_collected` field, and

-   save the cleaned tibble dataframe into a new tibble datatable called `mc3_nodes_cleaned`.

```{r}
mc3_nodes_cleaned <- mc3_nodes %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(id)) %>%
  distinct(id, .keep_all = TRUE) %>%
  select(-thing_collected)
```

### 4.2 Cleaning and wrangling edges

The code chunk below will be used to:

-   rename source and target fields to from_id and to_id respectively,

-   convert values in from_id and to_id fields to character data type,

-   exclude values in from_id and to_id which not found in the id field of mc3_nodes_cleaned,

-   exclude records whereby from_id and/or to_id values are missing, and

-   save the cleaned tibble dataframe and called it mc3_edges_cleaned.

```{r}
mc3_edges_cleaned <- mc3_edges %>%
  rename(from_id = source, 
         to_id = target) %>%
  mutate(across(c(from_id, to_id), 
                as.character)) %>%
  filter(from_id %in% mc3_nodes_cleaned$id, 
         to_id %in% mc3_nodes_cleaned$id) %>%
  filter(!is.na(from_id), !is.na(to_id))
```

The code chunk below will be used to create mapping of character id in `mc3_nodes_cleaned` to row index

```{r}
node_index_lookup <- mc3_nodes_cleaned %>%
  mutate(.row_id = row_number()) %>%
  select(id, .row_id)
```

The code chunk below will be used to join and convert `from_id` and `to_id` to integer indices. At the same time we also drop rows with unmatched nodes.

```{r}
mc3_edges_indexed <- mc3_edges_cleaned %>%
  left_join(node_index_lookup, 
            by = c("from_id" = "id")) %>%
  rename(from = .row_id) %>%
  left_join(node_index_lookup, 
            by = c("to_id" = "id")) %>%
  rename(to = .row_id) %>%
  select(from, to, is_inferred, type) %>%
  filter(!is.na(from) & !is.na(to))  
```

The code chunk below is used to subset nodes to only those referenced by edges.

```{r}
used_node_indices <- sort(
  unique(c(mc3_edges_indexed$from, 
           mc3_edges_indexed$to)))

mc3_nodes_final <- mc3_nodes_cleaned %>%
  slice(used_node_indices) %>%
  mutate(new_index = row_number())
```

The code chunk below will be used to rebuild lookup from old index to new index.

```{r}
old_to_new_index <- tibble(
  old_index = used_node_indices,
  new_index = seq_along(
    used_node_indices))
```

The code chunk below will be used to update edge indices to match new node table.

```{r}
mc3_edges_final <- mc3_edges_indexed %>%
  left_join(old_to_new_index, 
            by = c("from" = "old_index")) %>%
  rename(from_new = new_index) %>%
  left_join(old_to_new_index, 
            by = c("to" = "old_index")) %>%
  rename(to_new = new_index) %>%
  select(from = from_new, to = to_new, 
         is_inferred, type)
```

## 5 Building the tidygraph object

The code chunk below will be used to build the tidygraph object.

```{r}
mc3_graph <- tbl_graph(
  nodes = mc3_nodes_final,
  edges = mc3_edges_final,
  directed = TRUE
)
```

Examine the object by using str() after the tidygraph object is created.

```{r}
str(mc3_graph)
```

## 5 Visualising the knowledge graph

Several of the **ggraph** layouts involve randomisation. In order to ensure reproducibility, it is necessary to set the seed value before plotting by using the code chunk below.

```{r}
set.seed(1234)
```

In the code chunk below, ggraph functions are used to create the whole graph.

```{r}
ggraph(mc3_graph, 
       layout = "fr") +
  geom_edge_link(alpha = 0.3, 
                 colour = "gray") +
  geom_node_point(aes(color = `type`), 
                  size = 4) +
  geom_node_text(aes(label = type), 
                 repel = TRUE, 
                 size = 2.5) +
  theme_void()
```



